---
title: "Build your first phaserR game (hedgehog vignette)"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Build your first phaserR game (hedgehog vignette)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette walks through a minimal **phaserR + Shiny** game where a hedgehog:

- moves with arrow keys,
- plays animations,
- collects apples (overlap),
- collides with walls,
- avoids enemy sprites.

## 1) Basic app structure: `ui` and `server`

A phaserR game lives inside a regular Shiny app.

```{r eval=FALSE}
library(shiny)
library(phaserR)

game <- PhaserGame$new(width = 900, height = 550)

ui <- tagList(
  game$ui()
)

server <- function(input, output, session) {
  game$set_shiny_session()
}

shinyApp(ui, server)
```

## 2) Add static groups (background tiles and walls)

Static groups are perfect for scenery and obstacles that should not move.

```{r eval=FALSE}
# grass floor tiles
floor <- game$add_static_group(
  name = "floor",
  url = "assets/hedgehog/terrain/grass.png"
)

for (x in seq(100, 800, by = 200)) {
  for (y in seq(100, 500, by = 150)) {
    floor$create(x = x, y = y)
  }
}

# walls that player cannot pass through
walls <- game$add_static_group(
  name = "walls",
  url = "assets/hedgehog/terrain/wall.png"
)

walls$create(450, 80)
walls$create(450, 470)
walls$create(80, 280)
walls$create(820, 280)
```

## 3) Add sprite (the player hedgehog)

Create a player sprite from a sprite sheet.

```{r eval=FALSE}
hedgehog <- game$add_sprite(
  name = "hedgehog",
  url = "assets/hedgehog/sprites/hedgehog_32.png",
  x = 140,
  y = 260,
  frameWidth = 32,
  frameHeight = 32,
  frameCount = 5,
  frameRate = 6
)
```

## 4) Add player controls

Attach keyboard movement to the sprite.

```{r eval=FALSE}
hedgehog$add_player_controls(
  directions = c("left", "right", "up", "down"),
  speed = 220
)
```

## 5) Add animations

Add directional animations and play one as default.

```{r eval=FALSE}
moves <- c("move_left", "move_right", "move_up", "move_down")

for (move in moves) {
  hedgehog$add_animation(
    suffix = move,
    url = paste0("assets/hedgehog/sprites/hedgehog_", move, "_32.png"),
    frameWidth = 32,
    frameHeight = 32,
    frameRate = 5
  )
}

hedgehog$play_animation("hedgehog_move_right")
```

## 6) Add overlap with other objects (collect apples)

Use overlap when two objects can share space and trigger events.

```{r eval=FALSE}
score <- reactiveVal(0)
apples <- game$add_static_group("apples", "assets/hedgehog/perks/apple_20.png")

apples$create(260, 140)
apples$create(640, 180)
apples$create(730, 390)

score_text <- game$add_text(text = "Score: 0", id = "score", x = 20, y = 20)

game$add_overlap(
  sprite_name = "hedgehog",
  group_name = "apples",
  callback_fun = function(evt) {
    apples$disable(evt)        # hide collected apple
    score(score() + 1)
    score_text$set(paste("Score:", score()))
  },
  input = input
)
```

## 7) Add collision with other objects (walls)

Use collision when objects should block each other.

```{r eval=FALSE}
game$add_collider(
  sprite_name = "hedgehog",
  group_name = "walls"
)
```

## 8) Add enemy sprites

Create one or more enemies and move them. If enemy overlaps the player, end game.

```{r eval=FALSE}
enemy <- game$add_sprite(
  name = "badger_1",
  url = "assets/hedgehog/sprites/badger_move_left_50.png",
  x = 700,
  y = 300,
  frameWidth = 50,
  frameHeight = 50,
  frameCount = 1,
  frameRate = 1
)

enemy$set_in_motion(
  dirX = -1,
  dirY = 0,
  speed = 70,
  distance = 150,
  lag = 0
)

game$add_overlap(
  sprite_name = "hedgehog",
  object_two_name = "badger_1",
  callback_fun = function(evt) {
    showNotification("Game over: enemy got you!", type = "error")
  },
  input = input
)
```

## 9) Full minimal app

Put all pieces together:

```{r eval=FALSE}
library(shiny)
library(phaserR)

game <- PhaserGame$new(width = 900, height = 550)

ui <- tagList(game$ui())

server <- function(input, output, session) {
  game$set_shiny_session()

  floor <- game$add_static_group("floor", "assets/hedgehog/terrain/grass.png")
  for (x in seq(100, 800, by = 200)) for (y in seq(100, 500, by = 150)) floor$create(x, y)

  walls <- game$add_static_group("walls", "assets/hedgehog/terrain/wall.png")
  walls$create(450, 80); walls$create(450, 470); walls$create(80, 280); walls$create(820, 280)

  hedgehog <- game$add_sprite(
    name = "hedgehog", url = "assets/hedgehog/sprites/hedgehog_32.png",
    x = 140, y = 260, frameWidth = 32, frameHeight = 32, frameCount = 5, frameRate = 6
  )
  hedgehog$add_player_controls(c("left", "right", "up", "down"), speed = 220)

  apples <- game$add_static_group("apples", "assets/hedgehog/perks/apple_20.png")
  apples$create(260, 140); apples$create(640, 180); apples$create(730, 390)

  score <- reactiveVal(0)
  score_text <- game$add_text("Score: 0", id = "score", x = 20, y = 20)

  game$add_collider("hedgehog", group_name = "walls")

  game$add_overlap("hedgehog", group_name = "apples", callback_fun = function(evt) {
    apples$disable(evt)
    score(score() + 1)
    score_text$set(paste("Score:", score()))
  }, input = input)

  enemy <- game$add_sprite(
    "badger_1", "assets/hedgehog/sprites/badger_move_left_50.png",
    x = 700, y = 300, frameWidth = 50, frameHeight = 50, frameCount = 1, frameRate = 1
  )
  enemy$set_in_motion(dirX = -1, dirY = 0, speed = 70, distance = 150, lag = 0)

  game$add_overlap("hedgehog", object_two_name = "badger_1", callback_fun = function(evt) {
    showNotification("Game over: enemy got you!", type = "error")
  }, input = input)
}

shinyApp(ui, server)
```

## Next steps

- Add multiple enemies with random paths.
- Add level progression and win/lose states.
- Add boosts or power-ups with keyboard controls.
- Reuse patterns from `inst/examples/hedgehog.R` for a richer game loop.
